---
title: "Cow Milk Production Time Series"
author: 'Gamma: Stevyn Fessler, Sarah Franzen, Christopher Tsoi, Jianan Zhou'
date: "March 9, 2018"

---




```{r}
# Data steps:
# 
#   Transform:
#     a. Box-Cox
#     b. Log
#     c. Square root
# 
# 
# 
#   Differencing:
#     a. Seasonality
#     b. Trend
#   
#     check acf, pacf, plots and variance
# 
# 
# 
# 
#   
# Model Fitting:
#   a. Pure ar(p) using Yule-Walker
#   b. Pure ma(q) using MLE
# 
# Use loop to check the different models with arma(p,q)
# 
# Diagnostics
# 
# 
# Forecasting
```


```{r reading in csv and load packages, echo=False}
setwd("C:/Users/estaf/Documents/Pstat 174/Final_Project_PS_174/")
milk <-  read.csv(file = "monthly-milk-production-pounds-p.csv", header=TRUE)

# in csv file manually delete footer cells at the bottom to make reading data as time series easier


library(qpcR)
library(forecast)


# rename column of data
colnames(milk)[2] <- "Monthy.milk"


```









```{r convert to time series object}

# convert dataframe to time series object
milk.ts <- ts(data=milk[,2], start=c(1962,1), frequency = 12)

```






```{r plot, acf and pacf}

#Initial time series, acf and pacf plots
ts.plot(milk.ts,xlab="Months", ylab="Pounds of milk per cow per month")

# plot acf and pacf
acf(milk.ts, lag.max = 60)
pacf(milk.ts, lag.max = 36)




```


# Transform then difference

```{r transformations}

# Transformations


#log transform
milk.log <- log(milk.ts)



# box-cox transform
milk.boxcox <- boxcox(milk.ts ~ as.numeric(1:length(milk.ts)))


milk.lambda = milk.boxcox$x[which(milk.boxcox$y == max(milk.boxcox$y))]
milk.lambda

milk.bct <- (1/milk.lambda)*(milk.ts^milk.lambda -1)


# Box Cox gives lambda value close to 1/2, therefore use square root transform
milk.sq <- sqrt(milk.ts)



#compare transformation plots
ts.plot(milk.ts, main="Original")
ts.plot(milk.log, main="Log")
ts.plot(milk.sq, main="Square Root")
ts.plot(milk.bct, main="Box-Cox")



```
# Check again what box cox test is called and what lambda maximizes

Before we difference the data we test transformations on our original data. Beginning with the Box-Cox transformation we find the lambda which maximizes the normality of the data is 0.4646. This lambda is close to 0.5 so we will transform the data using the square root transformation by taking the square root of each observation. We also test the log transform of our data, but based on the value of lambda from our Box-Cox transformation we conclude that the square root is a better transformation.


```{r differencing square root}
ts.plot(milk.sq)
var(milk.sq)
acf(milk.sq, xlim=c(0,2.5))
pacf(milk.sq, xlim=c(0,2.5))


# difference for seasonality
milk.diff1.sq <- diff(milk.sq,lag=12, differences=1)

# plot deseasonalized data
ts.plot(milk.diff1.sq)
var(milk.diff1.sq)
acf(milk.diff1.sq, xlim=c(0,2.5), main="Deseasonalized")
pacf(milk.diff1.sq, xlim=c(0,2.5), main="Deseasonalized")



# difference for trend
milk.diff2.sq <- diff(milk.diff1.sq, differences=1)

# plot detrended data
ts.plot(milk.diff2.sq)
var(milk.diff2.sq)
acf(milk.diff2.sq, xlim=c(0,2.5), main="Detrended")
pacf(milk.diff2.sq, xlim=c(0,2.5), main="Detrended")

milk.diff3.sq <- diff(milk.diff2.sq, differences=1)

var(milk.diff3.sq)
# variance gets smaller each difference


# final data for square root transform
milk.sqf <- milk.diff2.sq

```










## *****Professor's choice for model fit*****
```{r model fitting for square root}
# fit an AR model using Yule Walker

fit.ar.sq <- ar(milk.sqf, method="yule-walker")
fit.ar.sq


# Plot residuals of AR model
plot(residuals(fit.ar.sq))

# Box Test for AR model
Box.test(residuals(fit.ar.sq),lag=13,type="Box-Pierce", fitdf=12)

# Shapiro test of AR model
shapiro.test(residuals(fit.ar.sq))
```


```{r model fitting loop square root }

library(qpcR)

# Calculate AICc for ARMA models with p and q running from 0 to 5
aiccs.milk.sq <- matrix(NA, nr = 6, nc = 6)
dimnames(aiccs.milk.sq) = list(p=0:5, q=0:5)



# loop for aic matrix
for(p in 0:5)
{
  for(q in 0:5)
  {
    aiccs.milk.sq[p+1,q+1] = AICc(arima(milk.sqf, order = c(p,0,q),method = "ML"))
  }
}
aiccs.milk.sq
(aiccs.milk.sq==min(aiccs.milk.sq))

# The lowest AIC model is arma(3,3)
# -128.35 = AIC

fit.arma.sq.3.3 <- arima(milk.sqf, order = c(3,0,3), method = "ML")
fit.arma.sq.3.3
```



#Diagnostics



```{r diagnostics check for square root differenced data}

#Shapiro-Wilk test for normality of errors
shapiro.test(residuals(fit.arma.sq.3.3))
hist(residuals(fit.arma.sq.3.3), breaks = 14)

# Box Tests for serial correlation test of errors
Box.test(residuals(fit.arma.sq.3.3), lag = 12 , type = "Box-Pierce", fitdf = 4)
Box.test(residuals(fit.arma.sq.3.3), lag = 13, type = "Ljung-Box", fitdf = 4)

# More Model fit tests
qqnorm(residuals(fit.arma.sq.3.3))
qqline(residuals(fit.arma.sq.3.3))



```




```{r differencing box cox transform}
# Check the variance
var(milk.bct)

# difference for seasonality
milk.diff1.bct <- diff(milk.bct,lag=12,1)
var(milk.diff1.bct)
acf(milk.diff1.bct)
pacf(milk.diff1.bct)

# add difference for trend
milk.diff2.bct <- diff(milk.diff1.bct,1)
var(milk.diff2.bct)



acf(milk.diff2.bct, lag.max = 100)
pacf(milk.diff2.bct, lag.max = 100)


# check difference for trend again
milk.diff3.bct <- diff(milk.diff2.bct,1)
var(milk.diff3.bct)


```

[1] 8.769204
[1] 0.2326241
[1] 0.07055325
[1] 0.1720925


We take our Box-Cox transformed data and try to difference it to remove any seasonality and trend that may be present before we build any models. Our transformed data starts out with variance of 8.769 and falls to 0.233 upon differencing for seasonality. We then difference once again for trend, and the variance drops again to 0.0706. We try differencing a third time and note the increase in variance to 0.1721. The smallest variance is given by the deseasonalized and once detrended data.




### TA suggested models, this one acutally works
### Spikes in pacf at 12 and 1 suggest ar component p=1 and seasoal ar component P=1
### Same thing for acf with q=1 and Q=1
sarima(1,0,1)x(1,1,1)12

First we will try to fit a mixed ARMA(p,q) model and a pure AR(p) model.


```{r fit model for box cox transform differenced}

# box cox transform yule walker model
fit.bct.yw <- ar(milk.diff2.bct, method="yule-walker")
fit.bct.yw
plot(residuals(fit.bct.yw))


shapiro.test(residuals(fit.bct.yw))

Box.test(residuals(fit.bct.yw), lag=13, type="Box-Pierce", fitdf = 12)

hist(residuals(fit.bct.yw), breaks = 15)

```
Similar to our other pure AR(12) model from the square transformed data, we have a high order AR model which does not make sense. It is not ideal to have a model with 12 predictors. This AR(12) model also fails the Shapiro-Wilk test for normality as well as the Box-Pierce test for independence. We will try to fit a mixed ARMA(p,q) model on our Box-Cox transformed and differenced data.


```{r model fitting loop for box cox}
#edit for our data, this is from lab 7

library(qpcR)




# Calculate AICc for ARMA models with p and q running from 0 to 5
aiccs.bct.milk.s <- matrix(NA, nr = 6, nc = 6)
dimnames(aiccs.bct.milk.s) = list(p=0:5, q=0:5)

for(p in 0:5)
{
  for(q in 0:5)
  {
    aiccs.bct.milk.s[p+1,q+1] = AICc(arima(milk.diff2.bct, order = c(p,0,q),method = "ML"))
  }
}
aiccs.bct.milk.s
(aiccs.bct.milk.s==min(aiccs.bct.milk.s))

fit.bct.3.12 <- arima(milk.diff2.bct, order = c(12, 0, 3), method = "ML", xreg=1 : length(milk.diff2.bct))

fit.bct.3.3 <- arima(milk.diff2.bct, order = c(3,0,3), method = "ML", xreg=1 : length(milk.diff2.bct))


# aic 15.44
fit.bct.3.3

qqnorm(residuals(fit.bct.3.3))
hist(residuals(fit.bct.3.3))
plot(residuals(fit.bct.3.3))
shapiro.test(residuals(fit.bct.3.3))
Box.test(residuals(fit.bct.3.3), lag = 12 , type = "Box-Pierce", fitdf = 4)


```


```{r box cox sarima plots for acf and pacf }


acf(milk.diff2.bct, lag.max = 100)
pacf(milk.diff2.bct, lag.max = 100)

```



```{r sarima fitting }

fit.sarima.bct <- arima(milk.bct, order = c(1,0,1), seasonal = list(order = c(1,1,1), period = 12), method = "ML")
fit.sarima.bct

shapiro.test(residuals(fit.sarima.bct))
Box.test(residuals(fit.sarima.bct), lag = 13, type="Box-Pierce", fitdf = 4)
Box.test(residuals(fit.sarima.bct), lag = 13, type="Ljung-Box", fitdf = 4)
Box.test( ((residuals(fit.sarima.bct))^2), lag=13, type="Ljung-Box", fitdf = 0)


acf(residuals(fit.sarima.bct), lag.max = 36)
pacf(residuals(fit.sarima.bct), lag.max = 36)

tsdiag(arima(residuals(fit.sarima.bct),order = c(0,0,0)))


qqnorm(residuals(fit.sarima.bct))
qqline(residuals(fit.sarima.bct))

hist(residuals(fit.sarima.bct), breaks = 15)
summary(fit.sarima.bct)


```


```{r forecasting}

library(forecast)
# Predict 10 future observations and plot
pred.bct <- predict(fit.sarima.bct, n.ahead=24)



milk.forecast <- forecast(fit.sarima.bct)

plot(milk.forecast)




pred.1977 <- sum(pred.bct$pred[1:12])
pred.1978 <- sum(pred.bct$pred[13:24])



fit.sarima.bct 
pred.bct



plot(1:length(milk.ts), milk.ts, main = "Monthly Milk per cow", ylab = "Milk in pounds per cow", xlab = "Date in Years", type = "l", xaxt = "null", xlim = c(130,193), ylim = c(750,1050))
points( (length(milk.ts)+1):(length(milk.ts)+24), ((milk.lambda*pred.bct$pred+1)^(1/milk.lambda)))


lines((length(milk.ts)+1):(length(milk.ts)+24),((milk.lambda*pred.bct$pred+1)^(1/milk.lambda) + 1.96*(milk.lambda*pred.bct$se+1)^(1/milk.lambda)),lty=2)

lines((length(milk.ts)+1):(length(milk.ts)+24),((milk.lambda*pred.bct$pred+1)^(1/milk.lambda) - 1.96*(milk.lambda*pred.bct$se+1)^(1/milk.lambda)),lty=2)


axis(side = 1, as.Date(seq(from = as.Date("1962-01-01"), to = as.Date("1977-12-31"), by = "month")))

```

